

---
title: "HR- Analytics"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    social: [ "twitter", "facebook", "menu"]
    source_code: embed
runtime: shiny
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(rsconnect)
#rsconnect::deployApp('mukeshvadapalli.shinyapps.io/HRAnalytics/')
library(flexdashboard)
library(shinythemes)
library(shiny)
library(dplyr)
library(ggplot2)
library(plotly)
library(tidyr)
library(forcats)
library(highcharter)
library(viridisLite)
library(ggcorrplot)
library(corrplot)
library(gridExtra)
library(grid)
```


```{r}
hranalytics <- read.csv('hranalytics.csv',head = TRUE,sep=",")
```


Dataset {data-icon="fas fa-database"}
=============================================================================

column {.sidebar}
----------------------------------------------------------------------------
**HR- Analytics**

Dataset consists of 1471 profiles of employees of a large company collected through an feedback survey. It involves different reasons that causes the attrition in the company. Attrition refers to the loss of staff by voluntary or involuntary reasons.
 
**What are the different factors that causes the attrition in the company.**

**Descibtion of important Variables in dataset:**

---------------------------------------------------------
- Attrition  Loss of staff from the company
- Department - Working departments in the company
- EducationField - Staff education background
- EnvironmentStatisfication - Statisfaction of working environment from 1 to 4
- JobStatisfaction - Statisfaction from job 1 to 4
- JobInvolvement - How seriously employ works in an organization score 1 to 4
- JobLevel - Position/ level in an organization
- RelationshipSatisfaction  - relationships with other working staff score 1 to 4
- WorkLifeBalance  - Balance between work and life 


```{r}
#hranalytics<-rename(hranalytics, c("ï..Age" = "ï..ï..Age"))
```


Row {data-height=550}
-----------------------------------------------------------------------

```{r}
DT::datatable(hranalytics,
              options = list(
                  pï..AgeLength = 25
              ))
```

Demographics
============================================================================
 <!-- {data-icon="fas fa-chart-bar"} -->
<!-- Demographics has ï..Age, Gender, MaritalStatus, Education and EducationField features -->


Row {data-height=450}
--------------------------------------------------------------------------

### Total number(percentï..Age) of male and female in company

```{r}
colors <- c('rgba(25, 181, 254, 1)', 'rgba(58, 83, 155, 1)')
mf<-hranalytics %>% 
  group_by(Gender) %>%
  summarize(no_of_employees=n(),avgï..Age=round(mean(ï..Age), 3))
fig <- plot_ly(mf, labels = ~Gender, values = ~no_of_employees, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        insidetextfont = list(color = '#FFFFFF'),
        hoverinfo = 'text',text = ~paste(Gender,avgï..Age),
        marker = list(colors = colors,
                      line = list(color = '#FFFFFF', width = 1)),
        showlegend = FALSE)
fig <- fig %>% layout(title = '% of Male and Female in the company',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

fig


```

<!-- Company has 60% of Male Staff and 40% of Female staff -->
 <!-- Averï..Age ï..Age of male is 36.653 and female is 37.33 -->

### AverAge salary by Gender

```{r}
genin <- hranalytics %>% select(Gender, MonthlyIncome) %>% group_by(Gender) %>% summarise(avg_income=round(mean(MonthlyIncome), 2)) %>%
ggplot(aes(x=Gender, y=avg_income)) + geom_bar(stat="identity", fill="#7abfff", width=0.5) + 
geom_text(aes(x=Gender, y=0.01, label= paste0("$ ", avg_income)),
            hjust=-2, vjust=0, size=3, 
            colour="black", fontface="bold",
         angle=360) + labs(title="Averï..Age Salary by Gender", x="Gender",y="Salary") + coord_flip() + 
theme_classic() + theme(plot.title=element_text(size=14, hjust=0.5))
genin
```

<!-- *Averï..Age income of female (6380.51)is higher than male(6686.57) in the company. -->

Row {data-height=400, .tabset}
--------------------------------------------------------------------------


### Employees from different Education Field

```{r}
Edufield<-hranalytics %>% 
  group_by(EducationField) %>%
  summarize(no_of_employees=n())
fig <- Edufield %>% plot_ly(labels = ~EducationField, values = ~no_of_employees)
fig <- fig %>% add_pie(hole = 0.5)
fig <- fig %>% layout(title = "Staff from different Education Field",  showlegend = T,
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

fig

```

<!-- * Company has 41.2% of total staffs from Life Sciences -->
<!-- * 31.6% are from the Medical background -->
<!-- * Lowest employees are from HR background -->



### Employees from different Educational Level
```{r}
hranalytics$Educational_Levels <-  ifelse(hranalytics$Education == 1, "Without College D.",
                            ifelse(hranalytics$Education == 2 , "Junior College D.",
                                  ifelse(hranalytics$Education == 3, "Bachelors D.",
                                        ifelse(hranalytics$Education == 4, "Masters D.", "Phd D."))))
hchart(hranalytics$Educational_Levels, colorByPoint = TRUE, name = "Educational_level")
```
<!-- * Educational level 1 is specified as 'without college' degree. -->
<!-- * Level 2 as with Junior college degree. -->
<!-- * Level 3 as Bachelors degree -->
<!-- * Level 4 as Phd degree. -->
<!-- * Company has large number of staff with Bachelor's level and less number from phd level. -->


ï..Age Income Educationlevel {data-icon="fas fa-users"}
================================================================================

Column {.sidebar}
-----------------------------------------------------------------------

```{r}
sliderInput(inputId = "ï..Agegroup",
            label = "ï..Age",
            min = min(hranalytics$ï..Age),
            max = max(hranalytics$ï..Age),
            value = c(20,50),
            step = 1)
```

```{r}


sliderInput(inputId = "Month",
            label = "MonthlyIncome",
            min = min(hranalytics$MonthlyIncome),
            max = max(hranalytics$MonthlyIncome),
            value = c(1009,5000),
            step = 1)
```
```{r}
checkboxGroupInput(inputId = "radio_input", 
             label = "Choose Education Level", 
             choices = unique(hranalytics$Educational_Levels), 
             selected = c("Masters D.","Bachelors D."))
```

-------------------------------------------------------------------------

Row {data-height=850}
-----------------------------------------------------------------------


### Monthly income according to ï..Age and Education level

```{r}
Montlydf <- reactive({
  hranalytics %>%
    filter(MonthlyIncome %>% between(left = input$Month[1],right = input$Month[2])) %>%
    filter(ï..Age %>% between(left = input$ï..Agegroup[1],right = input$ï..Agegroup[2]))%>%
    filter(Educational_Levels == input$radio_input)
})
```

```{r}
renderPlotly({
fig <- plot_ly(data = Montlydf(), x = ~ï..Age, y = ~MonthlyIncome, type = 'scatter',
  mode = 'markers', symbol = ~factor(Educational_Levels), symbols = c('circle','x'), marker = list(size = 10))

fig
})

```

<!-- * As seen from the above graph ï..Age is proportional to MonthlyIncome -->
<!-- * Older ï..Age groups are experienced soo there is increase in income, but it also depends on the educational level.  -->
<!-- * Educational level from 1 to 5 are spread through out the graph means educational degree is not much important as skillsets are more important. -->




Job roles
===========================================================================
<!-- {data-icon="fas fa-briefcase"} -->

Row {data-height=400} 
--------------------------------------------------------------------------

### Number of employees in a particular job roles
<!-- No of Employees in a particular Job Roles. -->

```{r}
hranalytics %>%
  group_by(JobRole) %>%
  summarize(no_of_employees=n()) %>%
hchart(type = "treemap",col = viridis(9),
       hcaes(x = JobRole, value= no_of_employees, color = no_of_employees)) %>% 
  hc_title(text = "No of Employees in each job role")
```


Row {data-height=400}
--------------------------------------------------------------------------

### Median salary according to job roles
<!-- Graphs shows the median salary of each Jobroles in the company. -->

```{r}
# Median Salary according to job roles
jobrsalary <- hranalytics %>% select(JobRole, MonthlyIncome) %>% group_by(JobRole) %>% summarize(med=median(MonthlyIncome))

fig <- plot_ly(jobrsalary, x = ~JobRole, y = ~med, type = 'bar',
        marker = list(color = c('rgba(30, 139, 195, 1)', 'rgba(0, 181, 204, 1)',
                                'rgba(44, 130, 201, 1)', 'rgba(1, 50, 67, 1)',
                                'rgba(25, 181, 254, 1)','rgba(89, 171, 227, 1)','rgba(82, 179, 217, 1)','rgba(34, 167, 240, 1)','rgba(52, 152, 219, 1)')))
fig <- fig %>% layout(title = "Median Salary of each Job Profile",
         xaxis = list(title = ""),
         yaxis = list(title = "Median Monthly Income"))

fig
```

 <!-- Manï..Agers and Research Directors have the highest salary on averï..Age. -->
 <!-- Laboratory Technician and Sales Representative have the lowest salary on averï..Age -->
 
### No of people working in each department by gender.

```{r}
# # How many people work in each department by gender
depgen <- hranalytics %>% group_by(Department, Gender) %>% summarise(No_of_Employees=n()) %>%
ggplot(aes(x=reorder(Department, -No_of_Employees), y=No_of_Employees, fill=Gender)) + geom_bar(stat="identity", position="dodge") + theme_minimal() + 
theme(axis.text.x = element_text(angle = 90), plot.title=element_text(hjust=0.5)) + scale_fill_manual(values=c("light blue", "Dodger Blue")) + 
labs(title="Number of Employees by Department",x="Department", y="Number of employees")

ggplotly(depgen)
```
<!-- * Research department has highest number of staff working of which males are more than 550 and females below 400. -->
<!-- * Human resources department has very low number of staff working. -->



Important Attrition Factor {data-icon="fa-user-times"}
==========================================================================


Row {data-width=500}
--------------------------------------------------------------------------




### Job roles and level of satisfaction with Attrition rate


```{r}
env.attr <- hranalytics %>% select(JobSatisfaction, JobRole, Attrition) %>% group_by(JobRole, Attrition) %>%
summarize(avgsat=mean(JobSatisfaction))

gdg <-ggplot(env.attr, aes(x=JobRole, y=avgsat)) + geom_line(aes(group=Attrition), color="#58ACFA", linetype="dashed") + 
geom_point(aes(color=Attrition), size=3)  + theme(axis.text.x=element_text(angle=90),panel.background = element_rect(fill = "white", colour = "grey50"),axis.line = element_line(size = 1, colour = "#000082")) +
labs(title="Job Satisfaction of each JobRole", y="Averï..Age Job Satisfaction", x="Job Position") 
ggplotly(gdg)
```

<!-- * Employees who rated high job satisfaction did not quit their job means no attrition. -->
<!-- * Healthcare Representatives rated high for job satisfaction but quit their job too . -->


### Attrition Rate of Department according to Averï..Age Income

```{r}
avg.income <- hranalytics %>% select(Department, MonthlyIncome, Attrition) %>% group_by(Attrition, Department) %>%
summarize(avg.inc=mean(MonthlyIncome)) %>%
ggplot(aes(x=reorder(Department, avg.inc), y=avg.inc, fill=Attrition)) + geom_bar(stat="identity", position="dodge") + facet_wrap(~Attrition) + 
theme_minimal() + theme(axis.text.x = element_text(angle = 90), 
                        plot.title=element_text(hjust=0.5),
                        panel.background = element_rect(fill = "white", colour = "grey50")) + 
scale_fill_manual(values=c("lightblue", "darkblue")) + 
labs(y="Averï..Age Income", x="Department", title="Averï..Age Income by Department \n and Attrition Status")


ggplotly(avg.income)
```

<!-- * There is a huge difference between averï..Age income of employees who quit and who didn't according to the department. -->
<!-- * Income plays an crucial role in attrition rate of the company -->

Row {data-width=550, .tabset}
--------------------------------------------------------------------------

### Attrition rate compared to total working year in this company and other company
```{r}
numworkyear <- hranalytics %>% filter(NumCompaniesWorked == 1) %>% 
  ggplot(aes(TotalWorkingYears, fill = Attrition)) + geom_density(alpha = 0.5) +
  ggtitle("Worked Only for this company") +  theme(plot.title = element_text(hjust = 0.5),
              panel.background = element_rect(fill = "white", colour = "darkblue"))
ggplotly(numworkyear)

```

<!-- The Attrition rate is highest in first 5 years of the employees who worked for this organization. -->


### Attrition due to overtime

```{r}
overtime_percent <- hranalytics %>% select(OverTime, Attrition) %>% filter(Attrition == "Yes") %>% group_by(Attrition, OverTime) %>%
summarize(n=n()) %>% mutate(pct=round(prop.table(n),2) * 100)

```

```{r}
colors <- c('#add8e6', 'rgba(58, 83, 155, 1)')
fig <- plot_ly(overtime_percent, labels = ~OverTime, values = ~n, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        insidetextfont = list(color = '#FFFFFF'),
        hoverinfo = 'text',text = ~paste(OverTime,n),
        marker = list(colors = colors,
                      line = list(color = '#FFFFFF', width = 1)),
        showlegend = TRUE)
fig <- fig %>% layout(title = '% of Attrition of staff due to OverTime',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

fig
```

<!-- * 53.6% of staff left the job due to OverTime -->
<!-- * Even due to working over time 46.4% of staff did not quit their job. -->
<!-- This shows that working overtime leads to attrition in the company. -->



### Attrition due to Distance from Home and Worklife Balance

```{r}
workdis<-ggplot(hranalytics, 
        aes(x= WorkLifeBalance, y=DistanceFromHome, group = WorkLifeBalance, fill = WorkLifeBalance)) + 
        geom_boxplot(alpha=0.7) + 
        theme(legend.position="none") + 
        facet_wrap(~ Attrition) + 
        ggtitle("Attrition due to Distance from Home and WorkLife Balance") + 
        theme(plot.title = element_text(hjust = 0.5),
              panel.background = element_rect(fill = "white", colour = "darkblue"))
ggplotly(workdis)
```

<!-- Employees who rated their worklife balance low were travelling from a farther away compared to those who rated their work-life balance as high. -->
<!-- Employees who quit their job were travelling from longer distance as compared to people still working. -->
<!-- Distance from home is one of the important factor for attrition. -->


Row {data-height=300}
--------------------------------------------------------------------------

### **Conclusion**


- **Attrition is mainly influenced by Averï..Age Income, working overtime, Distance from home, Job satisfaction.**
- **Employees who rated high satisfaction score didn't quit their job**
- **Employees having low averï..Age income in same department has quit their job compared to employees having high averï..Age income.**
- **The Attrition rate is highest in first 5 years of the employees who worked for this organization.**
- **As working years goes on increasing there is an decrease in Attrition**
- **About 53.6%, out of total employees who quit their job were due to working overtime**
- **Employees who rated their worklife balance low were travelling from a farther away compared to those who rated their work-life balance as high.**
- **Employees who quit their job were travelling from longer distance as compared to people still working.**

### **Submitted by**

- **Mukesh M Vadapalli**
- **C00246445**
- **Msc DataScience**
- **Prof -Dr Agnes Maciocha**
